---
title: "MP4"
author: "Meghan Suslovic & Yifan Ma"
date: "5/3/2018"
output:
  html_document:
    code_folding: hide
---

<<<<<<< HEAD

Context:
http://www.denofgeek.com/us/movies/female-directors/252810/26-female-directors-you-should-know-about-and-why-you-might-not 


-find percentage of female directors of feature length movies 
-is there a genre of movie that women direct in higher percentages
-how do the ratings differ
-are there more females working in a particular country (in info table)?
-how does their budget compare to movies with male directors?
-are females directing in higher ratios in independent films vs studio ones?
=======
>>>>>>> af54645542e9782205fd1549a265c3ab619c259d


```{r, message=FALSE}
library(mdsr)
library(tidyverse)
library(ggplot2)
library(RMySQL)
library(dplyr)
db <- dbConnect_scidb(dbname = "imdb")
```


##Where are all the women in Hollywood? Not behind the camera

Even without winning Best Director at the 90th Academy Awards for Lady Bird, Greta Gerwig still broke an eight year long steak. The steak she ended? Entirely male nominations for Best Director. The gap in time between female nominations is perhaps less shocking when the total number of women directing films is considered. The Center for the Study of Women in Television and Film found that, of the top 100 grossing films in 2017, only 8% of directors, 10% of writers, and 24% of producers were women. We were curious to see if these percentages remained the same for all feature length films and not just the top 100 grossing films. We used IMDb data to further examine the gender gap in the film industry. IMDb is an online database that maintains information on movies and tv series (among other categories) and all of our data was pulled from the online database.

### Directors

Gerwig’s nomination suggests there are very few female directors, but has this always been the case? Essentially, yes. In fact, the percent of female directors hit an all time high in 2017. As positive as that sounds, female directors are still only at 16%, so the lack of nominations for Best Director is not surprising. We limited our dataset to only include feature length films (40 minutes or longer) and the number of directors has steadily increased between 1911 and a little after the year 2000. From there it rapidly increases from 2,000 directors to just over 7,000 in 2014. Over this same time span, the percent of female directors has varied considerably more. Interestingly, the increase in percentage of female directors began around 1975, much earlier than the overall increase in number of directors. This ‘wave’ of female directors is likely a result of second-wave feminism, which picked up speed in the late 1960s.

```{r}
directors <- db %>%
  dbGetQuery("SELECT n. gender, mi. movie_id,t. Production_year, mi. info
FROM cast_info ci
JOIN role_type rt
ON ci. role_id = rt .id
JOIN name n
ON n.id =  ci. person_id
JOIN movie_info mi ON mi. movie_id = ci. movie_id
JOIN info_type it ON mi. Info_type_id = it.id
JOIN title t ON t.id  = ci. movie_id
JOIN kind_type kt ON kt. id = t. kind_id 
WHERE rt.id = 8 AND t. kind_id = 1 AND info_type_id = 1
HAVING info > 40
")

head(directors)

directors <- directors %>%
  filter (Production_year != 2018,Production_year != 2021, Production_year != 2024, Production_year != 2019, Production_year != 2020, !is.na(gender))
  

d_sum <- directors %>%
  group_by(Production_year) %>%
  summarise(sum_n = n())

d_timeline <- directors %>%
  filter(gender == "f") %>%
  group_by(Production_year) %>%
  summarise(f = n()) %>%
  left_join (d_sum, by ="Production_year") %>%
  mutate(p = f/sum_n*100)
  
d <- d_timeline %>%
  ggplot(aes( x = Production_year, y = p))+
  geom_line()+
  labs(title= "Percent of Female Directors (1911-2017) ", x = "Year of Production",  y = "Percent of Female Directors", caption = "Source: IMDB")+
  theme_linedraw()
  
t_total <- d_timeline %>%
  ggplot(aes( x = Production_year, y = sum_n))+
  geom_line()+
  labs(title= "Total Number of Directors (1911-2017) ", x = "Year of Production",  y = "Number of Directors", caption = "Source: IMDB")+
  theme_linedraw()

d
t_total  
```

### Directors by country

```{r}
directors_c <- db %>%
  dbGetQuery("SELECT n. gender, mi. movie_id,t. Production_year, mi. info
FROM cast_info ci
JOIN role_type rt
ON ci. role_id = rt .id
JOIN name n
ON n.id =  ci. person_id
JOIN movie_info mi ON mi. movie_id = ci. movie_id
JOIN info_type it ON mi. Info_type_id = it.id
JOIN title t ON t.id  = ci. movie_id
JOIN kind_type kt ON kt. id = t. kind_id 
WHERE rt.id = 8 AND t. kind_id = 1 AND  info_type_id = 8
") 

directors_c <- directors_c %>%
  filter (!is.na(gender), !is.na(Production_year), !is.na(info)) %>%
  right_join(directors, by = "movie_id")
  
directors_usa <- directors_c %>%
  filter (info.x == "USA") %>%
  select(gender.x, Production_year.x, info.x) %>%
  group_by(Production_year.x) %>%
  summarise(U.S = n())

  
directors_usaf <- directors_c %>%
  filter (info.x == "USA", gender.x == "f") %>%
  select(gender.x, Production_year.x, info.x) %>%
  group_by(Production_year.x) %>%
  summarise(f = n()) %>%
  left_join (directors_usa, by ="Production_year.x") %>%
  mutate(u.s = f/U.S*100)

#du <- directors_usaf %>%
 # ggplot(aes(x = Production_year.x, y = S)) +
  #geom_line()+
  #labs(title= "Total Number of Directors in the United States (1913-2017) ", x = "Year of Production",  y = "Number of Directors", caption = "Source: IMDB")+
  #theme_linedraw()

#duf <- directors_usaf %>%
 # ggplot(aes(x = Production_year.x, y = p)) +
  #geom_line()+
  #labs(title= "Percent of Female Directors in the United States (1913-2017) ", x = "Year of Production",  y = "Percent of Female Directors", caption = "Source: IMDB")+
  #theme_linedraw()

  
directors_o <- directors_c %>%
  filter(info.x != "USA" ) %>%
  select(gender.x, Production_year.x, info.x) %>%
  group_by(Production_year.x) %>%
  summarise(Other = n()) 
  
directors_of <- directors_c %>%
  filter(info.x != "USA", gender.x == "f" ) %>%
  select(gender.x, Production_year.x, info.x) %>%
  group_by(Production_year.x) %>%
  summarise(f = n()) %>%
  left_join (directors_o, by = "Production_year.x") %>%
  mutate(other = f/Other*100)

#do <- directors_of %>%
 # ggplot(aes(x = Production_year.x, y = S)) +
  #geom_line()+
  #labs(title= "Total Number of Directors in the Rest of the World (1911-2017) ", x = "Year of Production",  y = "Number of Directors", caption = "Source: IMDB")+
  #theme_linedraw()

#dof <- directors_of%>%
 # ggplot(aes(x = Production_year.x, y = p))+
  #geom_line()+
  #labs(title= "Percent of Female Directors in the Rest of the World (1911-2017) ", x = "Year of Production",  y = "Percent of Female Directors", caption = "Source: IMDB")+
  #theme_linedraw()

directors_a <- directors_o %>%
  inner_join(directors_usa, by = "Production_year.x") %>%
  ggplot(aes(x = Production_year.x, y = U.S))+
  geom_line(color = "red")+
  geom_line(aes(x = Production_year.x, y = Other),color = "blue")+
  labs(title= " Number of Directors in the Rest of the World VS. in the U.S. (1911-2017) ", x = "Year of Production",  y = "Number of Directors", caption = "Source: IMDB")+
  theme_linedraw()

directors_af <- directors_of %>%
  inner_join(directors_usaf, by = "Production_year.x") %>%
  ggplot(aes(x = Production_year.x, y = p.y))+
  geom_line(color = "red")+
  geom_line(aes(x = Production_year.x, y = p.x), color = "blue")+
  labs(title= "Percent of Female Directors in the Rest of the World VS. in the U.S. (1911-2017) ", x = "Year of Production",  y = "Percent of Female Directors", caption = "Source: IMDB")+
  theme_linedraw()

```

### Writers

```{r}
writer <- db %>%
  dbGetQuery("SELECT n. gender, mi. movie_id,t. Production_year, mi. info
FROM cast_info ci
JOIN role_type rt
ON ci. role_id = rt .id
JOIN name n
ON n.id =  ci. person_id
JOIN movie_info mi ON mi. movie_id = ci. movie_id
JOIN info_type it ON mi. Info_type_id = it.id
JOIN title t ON t.id  = ci. movie_id
JOIN kind_type kt ON kt. id = t. kind_id 
WHERE rt.id = 3 AND t. kind_id = 1 AND info_type_id = 1
HAVING info > 40
")

head(writer)

writer <- writer %>%
  filter (Production_year != 2018, Production_year != 2024, Production_year != 2019, Production_year != 2020, !is.na(gender))
  

w_sum <- writer %>%
  group_by(Production_year) %>%
  filter(Production_year != 1906, Production_year != 1908, Production_year != 1910) %>%
  summarise(sum_n = n())

w_timeline <- writer %>%
  filter(gender == "f") %>%
  group_by(Production_year) %>%
  summarise(f = n()) %>%
  left_join (w_sum, by ="Production_year") %>%
  mutate(p = f/sum_n*100)
  
w <- w_timeline %>%
  ggplot(aes( x = Production_year, y = p))+
  geom_line()+
  labs(title= "Percent of Female Writers (1912-2017) ", x = "Year of Production",  y = "Percent of Female Writers", caption = "Source: IMDB")+
  theme_linedraw()
  
w_total <- w_timeline %>%
  ggplot(aes( x = Production_year, y = sum_n))+
  geom_line()+
  labs(title= "Number of All Writers (1912-2017) ", x = "Year of Production",  y = "Number of Writers", caption = "Source: IMDB")+
  theme_linedraw()

w
w_total  
```

###Producers

```{r}
producers <- db %>%
  dbGetQuery("SELECT n. gender, mi. movie_id,t. Production_year, mi. info
FROM cast_info ci
JOIN role_type rt
ON ci. role_id = rt .id
JOIN name n
ON n.id =  ci. person_id
JOIN movie_info mi ON mi. movie_id = ci. movie_id
JOIN info_type it ON mi. Info_type_id = it.id
JOIN title t ON t.id  = ci. movie_id
JOIN kind_type kt ON kt. id = t. kind_id 
WHERE rt.id = 4 AND t. kind_id = 1 AND info_type_id = 1 HAVING info > 40
")

head(producers)

producers <- producers %>%
  filter (Production_year != 2018, Production_year != 2024, Production_year != 2019, Production_year != 2020, !is.na(gender))
  

p_sum <- producers %>%
  filter(Production_year != 1908, Production_year != 1910)%>%
  group_by(Production_year) %>%
  summarise(sum_n = n())

p_timeline <- producers %>%
  filter(gender == "f") %>%
  group_by(Production_year) %>%
  summarise(f = n()) %>%
  left_join (p_sum, by ="Production_year") %>%
  mutate(p = f/sum_n*100)
  
p <- p_timeline %>%
  ggplot(aes(x = Production_year, y = p))+
  geom_line()+
  labs(title= "Percent of Female Producers (1911-2017) ", x = "Year of Production",  y = "Percent of Female Producers", caption = "Source: IMDB")+
  theme_linedraw()
  
p_total <- p_sum %>%
  ggplot(aes( x = Production_year, y = sum_n))+
  geom_line()+
  labs(title= "Number of All Producers (1911-2017) ", x = "Year of Production",  y = "Number of Producers", caption = "Source: IMDB")+
  theme_linedraw()

p
p_total  
```

```{r}
genre <- db %>%
  dbGetQuery("SELECT n. gender, t. Production_year, mi.info
FROM cast_info ci
JOIN role_type rt
ON ci. role_id = rt .id
JOIN name n
ON n.id =  ci. person_id
JOIN movie_info mi ON mi. movie_id = ci. movie_id
JOIN info_type it ON mi. Info_type_id = it.id
JOIN title t ON t.id  = ci. movie_id
JOIN kind_type kt ON kt. id = t. kind_id 
WHERE  rt.id = 8 AND t. kind_id = 1 AND info_type_id = 3
")

```

##Genres 

```{r}
genre <- genre %>%
  select(gender, info) %>%
  filter (!is.na(gender), !is.na(info))

g_sum <- genre %>%
  group_by(info) %>%
  summarise(S = n())

genre_w <-genre %>%
  group_by(info) %>%
  summarise(f = sum(gender == "f"), m = sum(gender == "m"))%>%
  left_join(g_sum, by = "info") %>%
  mutate (f_w = f/S*100, m_w = m/S*100)
  

top_5f <- genre_w %>%
  arrange (desc(f_w)) %>%
  head(10) %>%
  ggplot(aes(x = info, y = f_w))+
  geom_bar(stat = "identity")


top_5f

top_5m <- genre_w %>%
  arrange (desc(m_w)) %>%
  head(5)%>%
  ggplot(aes(x = info, y = m_w))+
  geom_bar(stat = "identity")+
  labs(title= "Genres That Have the Most Male Directors", x = "Genre",  y = "Percent of Male Directors", caption = "Source: IMDB")+
 theme_linedraw()

Sum <-g_sum %>%
  head(10)%>%
  ggplot(aes (x = info, y = S)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x= element_text(angle=45)) +
  labs(title= "Number of Movies Produced in Each Genre", x = "Genre",  y = "Number of Movies", caption = "Source: IMDB")+
 theme_linedraw()



ptop_5 <- genre_w %>%
  arrange (desc(f_w)) %>%
  head(5) %>%
  ggplot(aes (x = info, y = f_w))+
  geom_bar(stat = "identity")+
  labs(title= "Genres That Have the Most Female Directors", x = "Genre",  y = "Percent of Female Directors ", caption = "Source: IMDB")+
  theme_linedraw()

Sum
ptop_5
```

According to the Motion Picture Association of America, 50% of moviegoers are women, suggesting the demand for movies directed from a female perspective is there. How can this vast underrepresentation of women in film be improved? Supporting movies with female directors by seeing the movie in theaters sends an economic message to the top production companies that we are interested in more female directed movies. An increase in female directors should also help reduce the gender gap in other aspects of movie production. The Center for the Study of Women in Television and Film found that films with at least one female director employed greater percentages of women working as writers, editors, cinematographers, and composers than films with all male directors. So go watch some female-directed movies!

Sources: 

[The Center for the Study of Women in Television and Film](https://womenintvfilm.sdsu.edu/wp-content/uploads/2018/01/2017_Celluloid_Ceiling_Report.pdf) 

[Encyclopaedia Britannica: The Silent Years, 1910-27](https://www.britannica.com/art/history-of-the-motion-picture/The-silent-years-1910-27)

[Women Film Pioneers Project: How Women Worked in the US Silent Film Industry](https://wfpp.cdrs.columbia.edu/essay/how-women-worked-in-the-us-silent-film-industry/)

[World Atlas: Largest Film Industries in the World](https://www.worldatlas.com/articles/largest-film-industries-in-the-world.html) 

[Motion Picture Association of America: 2017 Theme Report](https://www.mpaa.org/wp-content/uploads/2018/04/MPAA-THEME-Report-2017_Final.pdf) 

 
